---
output: html_document
title: "UKB results for pPGSxE proof-of-concept study"
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = F, message = F,warning = F, 
                      dpi = 150, fig.path = "../output/ukb_results/",
                      cache.path = "../cache/ukb_results/")
suppressMessages(silent <- lapply(
  c("knitr", "kableExtra", "tidyverse", "patchwork"), 
  library, character.only = TRUE))
theme_set(theme_bw())
```

```{r load-data}
ukb_subset_names <- c("training", "testing")
ukb_subset_names_clean <- c("Training", "Testing")
ukb_subsets <- map(set_names(ukb_subset_names), function(set) {
  read_csv(paste0("../data/processed/ukb_", set, "_set.csv"), show_col_types = FALSE) %>%
    mutate(across(matches("^gPC"), ~ . * bmi, .names = "bmiBy{.col}"))
})
ukb_phenos <- read_csv("../data/processed/ukb_phenos_unrelated.csv", show_col_types = FALSE) %>%
  mutate(across(matches("^gPC"), ~ . * bmi, .names = "bmiBy{.col}")) %>%
  filter(!is.na(bmi),
         !is.na(alt_log) | !is.na(ast_log) | !is.na(ggt_log))
ukb_subsets$full <- ukb_phenos

ancestry_df <- read_csv("../data/processed/ukb_phenos_panUKBB.csv", 
                        col_select = c(id, ancestry), show_col_types = FALSE)
ancestries <- unique(ancestry_df$ancestry[!is.na(ancestry_df$ancestry)])

ppgs_res_df_full <- readRDS("../data/processed/pgs/ppgs_res_df.rds") %>%
  mutate(pathway = gsub("HALLMARK_|KEGG_|KEGG_MEDICUS_", "", pathway)) %>%
  arrange(p.value)

ppgs_res_df <- ppgs_res_df_full %>%
  filter(term == "ppgs:bmi")

hallmark_annot_lines <- readLines("../data/processed/prsice/h.all.v2023.2.Hs.symbols.gmt")
hallmark_annot_df <- lapply(hallmark_annot_lines, function(line) {
  line_split <- strsplit(line, "\t")[[1]]
  pathway <- line_split[1]
  genes <- line_split[-(1:2)]
  tibble(pathway = pathway, gene = genes)
}) %>%
  bind_rows()

kegg_annot_lines <- readLines("../data/processed/prsice/c2.cp.kegg_legacy.v2024.1.Hs.symbols.gmt")
kegg_annot_df <- lapply(kegg_annot_lines, function(line) {
  line_split <- strsplit(line, "\t")[[1]]
  pathway <- line_split[1]
  genes <- line_split[-(1:2)]
  tibble(pathway = pathway, gene = genes)
}) %>%
  bind_rows()

kegg_medicus_annot_lines <- readLines("../data/processed/prsice/c2.cp.kegg_medicus.v2024.1.Hs.symbols.gmt")
kegg_medicus_annot_df <- lapply(kegg_medicus_annot_lines, function(line) {
  line_split <- strsplit(line, "\t")[[1]]
  pathway <- line_split[1]
  genes <- line_split[-(1:2)]
  tibble(pathway = pathway, gene = genes)
}) %>%
  bind_rows()

gene_pathway_annot_df <- bind_rows(list(
  hallmark = hallmark_annot_df, 
  kegg_legacy = kegg_annot_df,
  kegg_medicus = kegg_medicus_annot_df
), .id = "pathway_group")
```

```{r configuration}
exposures <- c("bmi")
exposures_clean <- c("BMI")

outcomes <- c("alt_log", "ast_log", "ggt_log")
outcomes_clean <- c("log(ALT)", "log(AST)", "log(GGT)")

pathway_groups <- c("kegg_legacy", "hallmark", "kegg_medicus")
pathway_groups_clean <- c("KEGG", "Hallmark", "KEGG Medicus")

thresholds <- c("5e-08", "0.001")
threshold_names <- paste0("Pt_", c("5e-08", "0.001"))
threshold_names_clean <- c("P<5e-8", "P<0.001")

ppgs_config_df <- expand_grid(
  e = exposures,
  y = outcomes,
  pathway_group = pathway_groups,
  threshold = thresholds
)
```

# Concept summary: pPGS for GxE

```{r workflow-diagram, eval=F}
knitr::include_graphics("../doc/workflow_diagram.pdf")
```

Overall dataset consists of unrelated, multi-ancestry individuals from the UKBB (N = `r nrow(ukb_phenos)`).

Primary metric for evaluation: significance of regression pPGSxE term

Variables being manipulated:

* Outcome: 3 liver-related biomarkers (ALT, AST, GGT)
* P&T threshold: 0.001, 5e-8

# Data distributions

Overall, `r nrow(ukb_phenos)` unrelated individuals from the UKB were used for analysis.

```{r histograms, fig.asp=0.4}
plot_histogram <- function(f) {
  ukb_subsets$testing %>%
    select(x = {{ f }}) %>%
    filter(!is.na(x)) %>%
    ggplot(aes(x = x)) +
    geom_histogram(stat = "bin", bins = 30) +
    labs(x = f)
}

alt_hist <- plot_histogram("alt_log")
ast_hist <- plot_histogram("ast_log")
ggt_hist <- plot_histogram("ggt_log")

alt_hist + ast_hist + ggt_hist

# for (outcome in outcomes) {
#   print(plot_histogram(outcome))
# }
```

```{r bmi-biomarker-correlations}
tibble(y = outcomes) %>%
  rowwise() %>%
  mutate(cor = list(broom::tidy(cor.test(ukb_phenos$bmi, ukb_phenos[[y]], use = "pairwise.complete.obs")))) %>%
  unnest(cor) %>%
  select(y, estimate, p.value) %>%
  kable(caption = "BMI main effects on biomarkers") %>%
  kable_styling(full_width = FALSE)
```

# Baseline: standard PGS (gwPGS) x E interaction

```{r pgs-metdata}
pgs_metadata_df <- tibble(y = outcomes) %>%
  rowwise() %>%
  mutate(metadata_tbl = list(read_tsv(paste0("../data/processed/pgs/", y, "_prset.prsice")))) %>%
  unnest(metadata_tbl) %>%
  select(y, Set, Threshold, n_snps = Num_SNP)

pgs_metadata_df %>%
  filter(Set == "Base") %>%
  kable(caption = "gwPGS metadata") %>%
  kable_styling(full_width = FALSE)
```

```{r plot-gwpgs-results}
ppgs_res_df %>%
  filter(pathway == "Base",
         pathway_group == "kegg_legacy") %>%  # Base is the same regardless of pathway group
  mutate(nlp = -log10(p.value),
         y = factor(y, levels = outcomes, labels = outcomes_clean)) %>%
  ggplot(aes(x = y, y = nlp, fill = threshold)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_manual(values = c("lightgray", "darkgray"), 
                    name = "P&T threshold") +
  labs(x = "", y = "-log10(p) for mPGSxBMI interaction")

ppgs_res_df %>%
  filter(pathway == "Base",
         threshold == 0.001,
         pathway_group == "kegg_legacy") %>%
  mutate(p.value = as.character(signif(p.value, 3))) %>%
  select(biomarker = y, pathway, pgs_p_threshold = threshold, estimate, std.error, p.value) %>%
  kable(caption = "Standard PGSxE results") %>%
  kable_styling()
```

Note: mPGS main effect optimal thresholds from our iPGS project UKB analysis were 0.0005 (ALT), 0.01 (AST), and 0.0001 (GGT).

Based on these results, we will use a threshold of 0.001 for the pPGSxE interaction analyses going forward.

# Pathway PGS interactions

For each pathway collection, significance is assigned based on a Bonferroni correction for the total number of pathways.

```{r ppgs-results}
n_hallmark <- length(unique(ppgs_res_df_full$pathway[ppgs_res_df_full$pathway_group == "hallmark"]))
n_kegg_legacy <- length(unique(ppgs_res_df_full$pathway[ppgs_res_df_full$pathway_group == "kegg_legacy"]))
n_kegg_medicus <- length(unique(ppgs_res_df_full$pathway[ppgs_res_df_full$pathway_group == "kegg_medicus"]))
bonferroni_list <- list(hallmark = 0.05 / n_hallmark, 
                        kegg_legacy = 0.05 / n_kegg_legacy, 
                        kegg_medicus = 0.05 / n_kegg_medicus)

sig_ppgs_res_df <- ppgs_res_df %>%
  rowwise() %>%
  mutate(bonferroni_thresh = bonferroni_list[[pathway_group]]) %>%
  ungroup() %>%
  filter(p.value < bonferroni_thresh)
```

```{r output-ppgs-interactions}
ppgs_res_df %>%
  arrange(p.value) %>%
  mutate(p.value = signif(p.value, 3)) %>%
  # select(biomarker = y, pathway, pgs_p_threshold = threshold, estimate, std.error, p.value) %>%
  select(Biomarker = y, Pathway = pathway, `Pathway collection` = pathway_group,
         `pPGS p-value threshold` = threshold,
         `pPGSxBMI estimate` = estimate, SE = std.error, P = p.value) %>%
  write_csv("../output/ppgs_interaction_estimates.csv")
```

## Hallmark pathways

P = `r n_hallmark` hallmark pathways.

```{r hallmark-results, fig.asp=1}
hallmark_gene_hist <- hallmark_annot_df %>%
  group_by(pathway) %>%
  summarise(n_genes = n(), .groups = "drop") %>%
  ggplot(aes(x = n_genes)) +
  geom_histogram() +
  labs(x = "# genes per pathway", y = "Pathway count")

hallmark_ppgs_res_plt <- ppgs_res_df %>%
  filter(pathway_group == "hallmark",
         threshold == 0.001,
         p.value < bonferroni_list$hallmark) %>%
  mutate(nlp = -log10(p.value),
         l95 = estimate - 1.96 * std.error,
         u95 = estimate + 1.96 * std.error) %>%
  arrange(estimate) %>%
  mutate(pathway = factor(pathway, levels = pathway, 
                          labels = str_replace(pathway, "HALLMARK_", ""))) %>%
  ggplot(aes(x = pathway, y = nlp)) +
  geom_hline(yintercept = 0, color = "gray") +
  geom_hline(yintercept = -log10(bonferroni_list$hallmark), 
             linetype = "dashed", color = "gray") +
  geom_bar(stat = "identity", width = 0.5, position = position_dodge(width = 0.5)) + 
  # geom_point(position = position_dodge(width = 0.5)) +
  # geom_errorbar(aes(ymin = l95, ymax = u95), width = 0, 
  #               position = position_dodge(width = 0.5)) +
  facet_wrap(vars(y), ncol = 1, scale = "free_y") +
  labs(x = "", y = "-log10(p-interaction)",
       title = "Significant pPGS interaction results: hallmark gene sets") +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9, size = 4))

hallmark_gene_hist / hallmark_ppgs_res_plt +
  plot_layout(heights = c(1, 3))

ppgs_res_df %>%
  filter(pathway_group == "hallmark",
         threshold == 0.001,
         pathway != "Base") %>%
  head(15) %>%
  mutate(p.value = as.character(signif(p.value, 3))) %>%
  select(biomarker = y, pathway, pgs_p_threshold = threshold, estimate, std.error, p.value) %>%
  kable(caption = "Top hallmark pathway interactions") %>%
  kable_styling()
```

## KEGG (legacy) pathways

P = `r n_kegg_legacy` KEGG pathways.

```{r kegg-legacy-results, fig.asp=1}
kegg_gene_hist <- kegg_annot_df %>%
  group_by(pathway) %>%
  summarise(n_genes = n(), .groups = "drop") %>%
  ggplot(aes(x = n_genes)) +
  geom_histogram() +
  labs(x = "# genes per pathway", y = "Pathway count")

kegg_ppgs_res_plt <- ppgs_res_df %>%
  filter(pathway_group == "kegg_legacy",
         threshold == 0.001,
         p.value < bonferroni_list$kegg_legacy) %>%
  mutate(nlp = -log10(p.value),
         l95 = estimate - 1.96 * std.error,
         u95 = estimate + 1.96 * std.error) %>%
  arrange(estimate) %>%
  mutate(pathway = factor(pathway, levels = unique(pathway))) %>%
  ggplot(aes(x = pathway, y = nlp)) +
  geom_hline(yintercept = 0, color = "gray") +
  geom_hline(yintercept = -log10(bonferroni_list$kegg_legacy), 
             linetype = "dashed", color = "gray") +
  geom_bar(stat = "identity", width = 0.5, position = position_dodge(width = 0.5)) + 
  facet_wrap(vars(y), ncol = 1, scales = "free_y") +
  labs(x = "", y = "-log10(p-interaction)",
       title = "Significant pPGS interaction results: KEGG pathways") +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9, size = 4))

kegg_gene_hist / kegg_ppgs_res_plt +
  plot_layout(heights = c(1, 3))

ppgs_res_df %>%
  filter(pathway_group == "kegg_legacy",
         threshold == 0.001,
         pathway != "Base") %>%
  head(15) %>%
  mutate(p.value = as.character(signif(p.value, 3))) %>%
  select(biomarker = y, pathway, pgs_p_threshold = threshold, estimate, std.error, p.value) %>%
  kable(caption = "Top KEGG pathway interactions") %>%
  kable_styling()
```

## KEGG (Medicus) pathways

P = `r n_kegg_medicus` KEGG Medicus pathways.

```{r kegg-medicus-results, fig.asp=1}
kegg_medicus_gene_hist <- kegg_medicus_annot_df %>%
  group_by(pathway) %>%
  summarise(n_genes = n(), .groups = "drop") %>%
  ggplot(aes(x = n_genes)) +
  geom_histogram() +
  labs(x = "# genes per pathway", y = "Pathway count")

kegg_medicus_ppgs_res_plt <- ppgs_res_df %>%
  filter(pathway_group == "kegg_medicus",
         threshold == 0.001,
         p.value < bonferroni_list$kegg_medicus) %>%
  mutate(nlp = -log10(p.value),
         l95 = estimate - 1.96 * std.error,
         u95 = estimate + 1.96 * std.error) %>%
  arrange(estimate) %>%
  mutate(pathway = factor(pathway, levels = unique(pathway))) %>%
  ggplot(aes(x = pathway, y = nlp)) +
  geom_hline(yintercept = 0, color = "gray") +
  geom_hline(yintercept = -log10(bonferroni_list$kegg_medicus), 
             linetype = "dashed", color = "gray") +
  geom_bar(stat = "identity", width = 0.5, position = position_dodge(width = 0.5)) + 
  facet_wrap(vars(y), ncol = 1, scales = "free_y") +
  labs(x = "", y = "-log10(p-interaction)",
       title = "Significant pPGS interaction results: KEGG medicus pathways") +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9, size = 4))

kegg_medicus_gene_hist / kegg_medicus_ppgs_res_plt +
  plot_layout(heights = c(1, 3))

ppgs_res_df %>%
  filter(pathway_group == "kegg_medicus",
         threshold == 0.001,
         pathway != "Base") %>%
  head(15) %>%
  mutate(p.value = as.character(signif(p.value, 3))) %>%
  select(biomarker = y, pathway, pgs_p_threshold = threshold, estimate, std.error, p.value) %>%
  kable(caption = "Top KEGG Medicus pathway interactions") %>%
  kable_styling()
```

## Pathway annotation "metadata"

```{r pathway-annotations-figure, fig.asp = 0.4}
gene_pathway_counts_df <- gene_pathway_annot_df %>%
  group_by(pathway_group, pathway) %>%
  summarise(n_genes = n(), .groups = "drop_last") %>%
  mutate(n_pathways = n()) %>%
  ungroup()
pathway_groups_with_counts <- gene_pathway_counts_df %>%
  distinct(pathway_group, n_pathways) %>%
  mutate(pathway_group = factor(pathway_group, levels = pathway_groups)) %>%
  arrange(pathway_group) %>%
  mutate(pathway_group_with_n = paste0(pathway_groups_clean[match(pathway_group, pathway_groups)],
                                      " (", n_pathways, " total)")) %>%
  pull(pathway_group_with_n)
gene_pathway_counts_df %>%
  mutate(pathway_group = factor(pathway_group, levels = pathway_groups, 
                                labels = pathway_groups_with_counts)) %>%
  ggplot(aes(x = n_genes)) +
  geom_histogram() +
  labs(x = "# genes per pathway", y = "Pathway count") +
  facet_wrap(vars(pathway_group), scales = "free_y")
```

## Overall pPGSxE counts

```{r overall-ppgsxe-counts}
sig_ppgs_counts_df <- ppgs_res_df %>%
  filter(threshold == 0.001,
         pathway != "Base") %>%
  rowwise() %>%
  mutate(sig_threshold = bonferroni_list[[pathway_group]]) %>%
  ungroup() %>%
  mutate(sig = p.value < sig_threshold) %>% 
  group_by(y, pathway_group) %>%
  summarise(n_sig = sum(sig), n_total = n(), .groups = "drop") 
sig_ppgs_counts_df %>%
  kable(caption = "pPGS findings by biomarker and pathway annotation") %>%
  kable_styling(full_width = FALSE)

sig_ppgs_res_df %>%
  filter(threshold == 0.001,
         pathway != "Base") %>%
  group_by(pathway_group) %>%
  distinct(pathway) %>%
  summarise(n_pathways = n(), .groups = "drop") %>%
  kable(caption = "# unique significant pathways per pathway collection (over all biomarkers)") %>%
  kable_styling(full_width = FALSE)

sig_ppgs_res_df %>%
  filter(threshold == 0.001,
         pathway != "Base") %>%
  group_by(y, pathway_group) %>%
  summarise(n_sig = n(), .groups = "drop") %>%
  mutate(y = factor(y, levels = outcomes, labels = outcomes_clean),
         pathway_group = factor(pathway_group, levels = pathway_groups, labels = pathway_groups_with_counts)) %>%
  ggplot(aes(x = y, y = n_sig, fill = pathway_group)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  scale_fill_brewer(palette = "Dark2", name = "") +
  labs(x = "", y = "# significant pPGSxBMI interactions")
```

```{r n-eff-discoveries-comparison}
sig_n_eff_df <- read_csv("../data/processed/pgs/sig_n_eff_tbl.csv")

sig_ppgs_counts_df %>%
  inner_join(sig_n_eff_df, by = c("y", "pathway_group")) %>%
  mutate(n_eff = round(n_eff, 1)) %>%
  select(y, pathway_group, `# pathways` = n_sig, `# effective pathways` = n_eff) %>%
  kable(caption = "# significant pathways: 'raw' versus effective") %>%
  kable_styling(full_width = FALSE)
```

```{r n-eff-threshold-comparison}
all_n_eff_df <- read_csv("../data/processed/pgs/all_n_eff_tbl.csv")

sig_comparison_df <- ppgs_res_df %>%
  rowwise() %>%
  mutate(std_bonferroni_thresh = bonferroni_list[[pathway_group]]) %>%
  ungroup() %>%
  inner_join(all_n_eff_df, by = c("y", "pathway_group")) %>%
  mutate(eff_bonferroni_thresh = 0.05 / n_eff) %>%
  mutate(sig_std = p.value < std_bonferroni_thresh,
         sig_eff = p.value < eff_bonferroni_thresh) %>%
  filter(threshold == 0.001,
         pathway != "Base") %>%
  group_by(y, pathway_group) %>%
  summarise(n_sig_std = sum(sig_std), 
            n_sig_eff = sum(sig_eff), 
            n_total = n(), 
            .groups = "drop")

sig_comparison_df %>%
  filter(pathway_group == "kegg_legacy") %>%
  select(y, `Pathway count-based threshold` = n_sig_std, `Effective pathway-based threshold` = n_sig_eff) %>%
  kable(caption = "Discovery of KEGG pathways using standard vs. effective Bonferroni correction") %>%
  kable_styling(full_width = FALSE)
```

## Was the P&T threshold important?

```{r pt-thresh-comparison, fig.asp=0.4}
ppgs_res_df %>%
  filter(pathway_group == "kegg_legacy") %>%
  select(y, threshold, pathway, statistic) %>%
  pivot_wider(names_from = "threshold", values_from = "statistic") %>%
  mutate(y = factor(y, levels = outcomes, labels = outcomes_clean)) %>% 
  ggplot(aes(x = `5e-08`, y = `0.001`)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  facet_wrap(vars(y)) +
  labs(x = "Z-statistic (p-threshold = P<5e-8)", 
       y = "Z-statistic (p-threshold = P<0.001)")
```

## Sensitivity analysis - separate training set

```{r held-out-testing}
ppgs_res_df_trainset <- readRDS("../data/processed/pgs/ppgs_res_df_trainset.rds") %>%
  mutate(pathway = gsub("HALLMARK_|KEGG_|KEGG_MEDICUS_", "", pathway)) %>%
  arrange(p.value)

trainset_plt_df <- ppgs_res_df_trainset %>%
  filter(term == "ppgs:bmi") %>%
  mutate(nlp = -log10(p.value)) %>%
  select(source, pathway, statistic) %>%
  pivot_wider(names_from = "source", values_from = "statistic") %>%
  mutate(held_out_adj = held_out * sqrt(100 / 70)) 

trainset_plt <- trainset_plt_df %>%
  ggplot(aes(x = full, y = held_out)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  labs(x = "Z-statistic (100/20)", 
       y = "Z-statistic (80/20)")

trainset_plt

# adj_trainset_plt <- trainset_plt_df %>%
#   ggplot(aes(x = full, y = held_out_adj)) +
#   geom_point() +
#   geom_abline(slope = 1, intercept = 0) +
#   labs(x = "Z-statistic (100/20)", 
#        y = "Z-statistic (80/20) - sample size-adjusted")
# 
# trainset_plt + adj_trainset_plt +
#   plot_layout(ncol = 2) +
#   plot_annotation(tag_levels = "a")
# 
# lapply(trainset_plt_df[, 2:4], function(x) summary(abs(x))) %>%
#   setNames(c("held_out", "full", "held_out_adj"))
```

```{r bmi-pgs, fig.asp=0.4}
ppgs_res_df_bmiPGS <- readRDS("../data/processed/pgs/ppgs_res_df_bmiPGS.rds") %>%
  mutate(pathway = gsub("HALLMARK_|KEGG_|KEGG_MEDICUS_", "", pathway),
         adjustment = factor(adjustment, levels = c("bmi_pgs_main", "bmi_pgs_by_bmi"))) %>%
  arrange(p.value)

strip_labels <- c(
  bmi_pgs_main   = 'Adj.~PGS[BMI]',
  bmi_pgs_by_bmi = 'Adj.~PGS[BMI]~"×"~BMI'
)

ppgs_res_df %>%
  filter(y == "alt_log",
         pathway_group == "kegg_legacy",
         threshold == 0.001,
         term == "ppgs:bmi") %>%
  left_join(ppgs_res_df_bmiPGS, by = c("e", "y", "pathway_group", "threshold", 
                                       "pathway", "term"),
            suffix = c(".primary", ".bmiPgsAdj")) %>%
  # mutate(nlp = -log10(p.value)) %>%
  # select(source, pathway, statistic) %>%
  # pivot_wider(names_from = "source", values_from = "statistic") %>%
  ggplot(aes(x = statistic.primary, y = statistic.bmiPgsAdj)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0) +
  labs(x = "Z-statistic (primary model)", 
       y = "Z-statistic (adjusted model)") +
    facet_grid(
      cols     = vars(adjustment),
      labeller = as_labeller(strip_labels, label_parsed)
    )


bmi_pgs_df <- read_table("../data/processed/pgs/bmi_prset_hallmark.all_score", show_col_types = FALSE) %>%
  select(id = IID, bmi_pgs = `Base_0.001`)
test_ppgs_df <- read_csv("test_ppgs_subset.csv")
a <- ukb_phenos %>%
  inner_join(bmi_pgs_df, by = "id") %>%
  inner_join(test_ppgs_df, by = "id")

# lm(alt_log ~  bmi * GM, data = a) %>%
#   broom::tidy()
# lm(alt_log ~ bmi * bmi_pgs + bmi * GM, data = a) %>%
#   broom::tidy()
```

# Variant-specific GWIS

```{r load-gwis}
gwis_res_nom_list <- lapply(set_names(outcomes), function(y) {
  read_tsv(paste0("../data/processed/gwas/bmi_", y, "_merged_clean_nom"),
           show_col_types = FALSE)
})

gwis_bonferroni <- 5e-8
```

```{r gwis-manhattan, fig.width=8}
mh_data <- bind_rows(gwis_res_nom_list, .id = "y")  %>%
  mutate(P = P_int) %>%
  filter(!is.na(P)) %>%
  mutate(nlp = -log10(P))

# Trim points in crowded regions (credit to RaMWAS package for code snippet)
yfac <- as.integer(mh_data$nlp * 100) + 1L
yorder <- sort.list(yfac)
yfac <- factor(yfac, levels = as.character(seq_len(max(yfac))))
ygroup <- split(seq_along(yfac), yfac)
for (i in seq_along(ygroup)) {
  if (length(ygroup[[i]]) > 300) {
    ygroup[[i]] <- sample(ygroup[[i]], size = 300, replace = FALSE)
  }
}
keep <- unlist(ygroup, use.names=FALSE)

mh_data <- mh_data %>%
  select(SNP, CHR, POS, nlp, y) %>%
  dplyr::slice(keep) %>%
  mutate(POS = as.numeric(as.character(POS)),
         CHR = factor(CHR, levels = 1:22)) %>%
  arrange(CHR, POS) %>%
  mutate(pos_idx = seq(1, nrow(.)))

chr_lengths <- sapply(1:22, function(chr) with(mh_data, max(POS[CHR == chr])))
chr_start_pos <- cumsum(chr_lengths) - chr_lengths

mh_data <- mh_data %>%
  mutate(x_coord = chr_start_pos[CHR] + POS,
         color = ifelse(nlp > -log10(gwis_bonferroni), y, CHR),
  ) %>%
  arrange(as.integer(color), nlp)

lims <- mh_data %>%
  group_by(CHR) %>%
  summarise(avg_coord = (min(x_coord) + max(x_coord)) / 2)

mh_cols <- setNames(
  rep(x=c("#999999", "#555555"), length.out = 22),  # Gray/dark gray for alternating chromosomes
  levels(factor(lims$CHR))
)

mh_cols <- c(
  setNames(rep(c("#999999", "#555555"), length.out = 22), levels(factor(lims$CHR))),  # Gray/dark gray for chromosomes
  setNames(RColorBrewer::brewer.pal(3, "Set1"), unique(mh_data$y))   # Unique colors for `y`
)

gwis_mh_plt <- ggplot() +
  geom_point(data = mh_data,
             aes(x = x_coord, y = nlp, color = factor(color)), 
             size = 0.6, alpha = 1) +
  geom_hline(yintercept = -log10(gwis_bonferroni), linetype = "dashed", color = "black") + 
  scale_x_continuous(breaks=lims$avg_coord[c(1:16, 18, 20, 20, 22)], 
                     labels=c(1:16, 18, 20, 20, 22)) +
  scale_y_continuous(name=expression(-log[10](italic(p))*" interaction")) +
  scale_colour_manual(values = mh_cols, breaks = unique(mh_data$y),
                      name = "Biomarker") +
  theme_minimal() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(vjust = -1.5),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank())

(
  gwis_mh_plt + 
    scale_y_continuous(breaks = c(0, 10, 20, 30, 140)) +
    ggbreak::scale_y_break(c(30, 140))
) +
  theme(panel.border = element_blank(), axis.text.y.right = element_blank())
```

```{r pruning}
prune_chromosome <- function(chr_df, pval_col, locus_width) {
  # Prune variants given a chromosome-specific summary statistic data frame
  df <- arrange(chr_df, !!sym(pval_col))  # Sort by ascending p-value
  pruned_df <- tibble()
  while(nrow(df) > 0) {
    pruned_df <- bind_rows(pruned_df, df[1, ])  # Add lowest p-value to pruned dataset
    df <- filter(df, (POS < df$POS[1] - locus_width / 2) |  # Remove rest of variants in that distance-based locus
                   (POS > df$POS[1] + locus_width / 2))
  }
  pruned_df
}

prune_variants <- function(ss_df, pval_col, thresh = 5e-8, locus_width = 1e6) {
  # Prune variants across all chromosomes using a simple distance-based approach
  if (min(ss_df[[pval_col]], na.rm = TRUE) > thresh) return(tibble())
  ss_df %>%
    filter(!!sym(pval_col) < thresh) %>%
    mutate(POS = as.numeric(POS)) %>%
    nest(data = -CHR) %>%
    mutate(pruned_ss = map(data, prune_chromosome, pval_col, locus_width)) %>%
    unnest(pruned_ss) %>%
    select(-data) %>%
    dplyr::rename(index_var = SNP)
}

pruned_gwis_ss <- lapply(gwis_res_nom_list, prune_variants, "robust_P_int")
```

```{r gwis-outputs}
all_top_gwis_df <- pruned_gwis_ss %>%
  bind_rows(.id = "biomarker")

all_top_gwis_df %>%
  select(`Index variant` = index_var, CHR, POS, 
         `Non-effect allele` = NEA, `Effect allele` = EA, `Effect allele frequency` = AF, 
         `Beta-interaction` = `Beta_G-bmi`, `P-interaction` = robust_P_int, 
         `Beta-marginal (G)` = Beta_Marginal, `P-marginal (G)` = robust_P_marg) %>%
  write_csv("../output/top_gwis_loci.csv")

all_top_gwis_counts_df <- all_top_gwis_df %>%
  group_by(biomarker) %>%
  summarise(N_loci = n())

all_top_gwis_counts_df %>%
  kable(caption = "# GWIS hits per biomarker")
```

# Comparison of pPGS and gwPGS approaches (variance explained)

```{r lrt}
lrt_res_df <- expand_grid(
  y = outcomes
  # type = c("gw", "pwy")
) %>%
  rowwise() %>%
  mutate(lrt = list(readRDS(paste0("../data/processed/pgs/", 
                              y, "_kegg_legacy_pwy_vs_base_lrt.rds"))),
         lrt_p = lrt[["Pr(>Chisq)"]][2]) %>%
  ungroup()

lrt_res_df %>%
  select(y, lrt_p) %>%
  mutate(lrt_p = format(lrt_p, digits = 3)) %>%
  kable(caption = "P-values from LRT adding all significant pPGSxBMI to a gwPGSxBMI model") %>%
  kable_styling(full_width = FALSE)
```

# Comparison of pPGS and GWIS approaches (locus identification)

```{r snp-to-pathway-mapping}
gtf_df <- read_tsv("../data/processed/prsice/Homo_sapiens.GRCh37.75.gtf.gz",
                   skip = 5, col_names = c("CHR", "source", "type", "start", "stop",
                                           paste0("V", 1:3), "metadata")) %>%
  filter(type == "gene") %>%
  mutate(gene = str_match(metadata, 'gene_name "([^"]+)"')[,2]) %>%
  select(gene, CHR, tss = start)

locus_to_gene_df <- bind_rows(pruned_gwis_ss, .id = "y") %>%
  select(y, index_var, CHR, POS, robust_P_int) %>%
  mutate(CHR = as.character(CHR)) %>%
  left_join(gtf_df, by = "CHR") %>%
  mutate(distance = abs(tss - POS)) %>%
  filter(distance < 100000) %>%
  select(y, index_var, CHR, variant_P_int = robust_P_int, gene)
gene_to_sig_pathway_df <- gene_pathway_annot_df %>%
  mutate(pathway = gsub("HALLMARK_|KEGG_|KEGG_MEDICUS_", "", pathway)) %>%
  inner_join(filter(sig_ppgs_res_df, threshold == 0.001, pathway != "Base"), 
             by = c("pathway", "pathway_group")) %>%
  select(pathway_group, pathway, y, gene, pathway_P_int = p.value)

locus_pathway_mapping_df <- full_join(
  locus_to_gene_df,
  gene_to_sig_pathway_df,
  by = c("y", "gene")
)
# locus_pathway_mapping_df %>%
#   write_csv("../output/locus_to_pathway_table.csv")

locus_to_kegg_pathway_df <- left_join(
  locus_to_gene_df,
  filter(gene_to_sig_pathway_df, pathway_group == "kegg_legacy"),
  by = c("y", "gene")
) %>%
  filter(!is.na(index_var)) %>%
  group_by(y, index_var) %>%
  slice_min(pathway_P_int, with_ties = TRUE) %>%
  rename(top_pathway = pathway) %>%
  group_by(y, index_var, CHR, variant_P_int, top_pathway, pathway_group, pathway_P_int) %>%
  summarise(genes = ifelse(any(!is.na(top_pathway)), 
                           paste(gene, collapse = ", "),
                           NA)) %>%
  arrange(variant_P_int)
locus_to_kegg_pathway_df %>%
  filter(y == "alt_log") %>%
  kable(caption = "GWIS locus -> pathway mapping (ALT)") %>%
  kable_styling(full_width = FALSE)

pathway_to_locus_df <- left_join(
  gene_to_sig_pathway_df,
  locus_to_gene_df,
  by = c("y", "gene")
) %>%
  filter(!is.na(pathway)) %>%
  group_by(y, pathway) %>%
  slice_min(variant_P_int, with_ties = TRUE) %>%
  rename(top_index_var = index_var) %>%
  group_by(y, pathway, pathway_group, pathway_P_int, top_index_var, CHR, variant_P_int) %>%
  summarise(genes = ifelse(any(!is.na(top_index_var)), 
                           paste(gene, collapse = ", "),
                           NA), 
            .groups = "drop") %>%
  arrange(pathway_P_int)
pathway_to_locus_df %>%
  filter(y == "alt_log",
         pathway_group == "kegg_legacy") %>%
  kable(caption = "Pathway -> GWIS locus mapping (ALT)") %>%
  kable_styling(full_width = FALSE)
```

```{r gwis-vs-ppgs}
gwis_vs_ppgs_df <- pathway_to_locus_df %>%
  # filter(threshold == 0.001) %>%
  # select(y, pathway_group, pathway, estimate, std.error, p.value) %>%
  # left_join(select(locus_to_pathway_df, y, pathway, index_var), 
  #           by = c("y", "pathway")) %>%
  group_by(y, pathway_group) %>%
  summarise(n_pathways_without_gwis = sum(is.na(top_index_var)),
            n_pathways = n(), 
            .groups = "drop") %>%
  mutate(frac_unexplained = round((n_pathways_without_gwis / n_pathways), 2))

gwis_vs_ppgs_tbl <- gwis_vs_ppgs_df %>%
  filter(pathway_group == "kegg_legacy") %>%
  inner_join(all_top_gwis_counts_df, by = c("y" = "biomarker")) %>%
  mutate(y = factor(y, levels = outcomes, labels = outcomes_clean),
         pct_unexplained = paste0(round(100 * frac_unexplained), "%")) %>%
  mutate(across(everything(), as.character)) %>%
  select(Biomarker = y, `# GWIS loci` = N_loci, `# pPGSxBMI` = n_pathways, 
         `# "unexplained" pPGSxBMI pathways` = n_pathways_without_gwis, 
         `Fraction "unexplained" pPGSxBMI pathways` = pct_unexplained)

gwis_vs_ppgs_tbl %>%
  write_csv("../output/gwis_vs_ppgs_table.csv")

gwis_vs_ppgs_tbl %>%
  # pivot_longer(-y, names_to = "quantity", values_to = "value") %>%
  # pivot_wider(names_from = y, values_from = value) %>%
  kable(caption = "Are significant pPGS findings 'explained' by GWIS loci?") %>%
  kable_styling()
```

## Could we have tested for enrichment of GWIS instead?

```{r magma}
magma_res_df <- map(outcomes, function(y) {
  read_table(paste0("../data/processed/magma/bmi_", y, ".gsa.out"), skip = 3)
}) %>%
    setNames(outcomes) %>%
  bind_rows(.id = "y") %>%
  select(y, pathway = FULL_NAME, p_magma = P)

magma_ppgs_comparison_df <- magma_res_df %>%
  mutate(pathway = gsub("KEGG_", "", pathway)) %>%
  inner_join(filter(ppgs_res_df, pathway_group == "kegg_legacy"), 
             by = c("y", "pathway")) %>%
  filter(threshold == 0.001) %>%
  mutate(nlp_ppgs = -log10(p.value),
         nlp_magma = -log10(p_magma))

magma_ppgs_comparison_df %>%
  mutate(y = factor(y, levels = outcomes, labels = outcomes_clean)) %>%
  ggplot(aes(x = nlp_ppgs, y = nlp_magma)) +
  geom_hline(yintercept = 0, color = "gray") +
  geom_vline(xintercept = 0, color = "gray") +
  geom_point() +
  geom_abline() +
  # geom_smooth(method = "lm", se = FALSE) +
  labs(x = expression(-log[10](italic(p))*" pPGSxBMI"),
       y = expression(-log[10](italic(p))*" MAGMA")) +
  facet_wrap(vars(y), nrow = 1, scales = "free_x") +
  coord_cartesian(xlim = c(0, 10))

magma_ppgs_nlp_corr = cor.test(magma_ppgs_comparison_df$nlp_ppgs, 
                               magma_ppgs_comparison_df$nlp_magma,
                               method = "spearman")

magma_ppgs_comparison_df %>%
  filter(pathway == "GLYCEROLIPID_METABOLISM") %>%
  select(y, pathway, p_magma, p_ppgs = p.value)
```

The Spearman correlation between -log(p) from pPGSxBMI versus MAGMA-based GWIS enrichment is `r round(magma_ppgs_nlp_corr$estimate, 2)`.

# Specific visualizations

All using p-threshold = 0.001.

```{r base-pgs-plot-function}
generate_base_plot <- function(y, y_name) {
  base_sumstats <- ppgs_res_df_full %>%
    filter(pathway_group == "kegg_legacy", 
           y == .env$y,
           pathway == "Base",
           threshold == 0.001)
  base_ppgs_beta_main <- base_sumstats$estimate[base_sumstats$term == "ppgs"]
  base_ppgs_beta_int <- base_sumstats$estimate[base_sumstats$term == "ppgs:bmi"]
  
  base_pgs_plt <- tibble(bmi_grp = seq(-2, 2),
                         total_ppgs_beta = base_ppgs_beta_main + bmi_grp * base_ppgs_beta_int) %>%
    mutate(bmi_grp = factor(bmi_grp, labels = c("-2 SD", "-1 SD", "Mean", "+1 SD", "+2 SD"))) %>%
    ggplot(aes(x = bmi_grp, y = total_ppgs_beta, group = 1)) +
    geom_point() +
    geom_line() +
    geom_hline(yintercept = 0, color = "gray") +
    coord_cartesian(ylim = c(0, NA)) +
    labs(x = "BMI", y = paste0("Full PGS effect on ", y_name))
  
  base_pgs_plt
}
```

```{r ppgs-plot-function}
generate_ppgs_plot <- function(y, y_name) {
  ppgs_plt_df <- ppgs_res_df %>%
    filter(y == .env$y,
           pathway_group == "kegg_legacy",
           threshold == 0.001,
           # pathway != "Base",
           p.value < bonferroni_list[["kegg_legacy"]]) %>%
    mutate(base = factor(pathway == "Base", labels = c("pPGS", "Standard PGS")),
           pathway = str_replace(pathway, "Base", "Standard PGS"),
           pathway = str_trunc(pathway, 20, side = "right", ellipsis = "_"),
           pathway = factor(pathway, levels = unique(pathway)),
           nlp = -log10(p.value))
  ppgs_plt <- ppgs_plt_df %>%
    ggplot(aes(x = fct_rev(pathway), y = nlp)) +
    # geom_hline(yintercept = 0, color = "gray") +
    # geom_point(position = position_dodge(width = 0.5)) +
    # geom_errorbar(aes(ymin = l95, ymax = u95), width = 0, 
    #               position = position_dodge(width = 0.5)) +
    geom_bar(stat = "identity") +
    geom_hline(yintercept = -log10(bonferroni_list[["kegg_legacy"]]), 
               linetype = "dashed", color = "gray") +
    labs(x = "", y = "Std. pPGSxBMI effect (95% CI)") +
    labs(x = "", y = expression(-log[10](italic(p))*" pPGSxBMI")) +
    facet_grid(cols = vars(fct_rev(base)), scales = "free", space = "free") +
    theme(axis.text.x = element_text(angle = 45, hjust = 0.9, size = 4),
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank(),
          strip.text = element_blank(), strip.background = element_blank())
    # annotate("label", label = "KEGG pathways", 
    #          x = 15, y = 0.8 * max(ppgs_plt_df$nlp), size = 3)
  
  ppgs_plt
}
```

```{r manhattan-plot-function}
generate_gwis_plot <- function(y, y_name) {
  mh_data <- gwis_res_nom_list[[y]] %>%
    mutate(y = .env$y,
           P = robust_P_int) %>%
    filter(!is.na(P)) %>%
    mutate(nlp = -log10(P))
  
  # Trim points in crowded regions (credit to RaMWAS package for code snippet)
  yfac <- as.integer(mh_data$nlp * 100) + 1L
  yorder <- sort.list(yfac)
  yfac <- factor(yfac, levels = as.character(seq_len(max(yfac))))
  ygroup <- split(seq_along(yfac), yfac)
  for (i in seq_along(ygroup)) {
    if (length(ygroup[[i]]) > 300) {
      ygroup[[i]] <- sample(ygroup[[i]], size = 300, replace = FALSE)
    }
  }
  keep <- unlist(ygroup, use.names=FALSE)
  
  mh_data <- mh_data %>%
    select(SNP, CHR, POS, nlp, y) %>%
    dplyr::slice(keep) %>%
    mutate(POS = as.numeric(as.character(POS)),
           CHR = factor(CHR, levels = 1:22)) %>%
    arrange(CHR, POS) %>%
    mutate(pos_idx = seq(1, nrow(.)))
  
  chr_lengths <- sapply(1:22, function(chr) with(mh_data, max(POS[CHR == chr])))
  chr_start_pos <- cumsum(chr_lengths) - chr_lengths
  
  mh_data <- mh_data %>%
    mutate(x_coord = chr_start_pos[CHR] + POS,
           color = ifelse(nlp > -log10(gwis_bonferroni), y, CHR),
    ) %>%
    arrange(as.integer(color), nlp)
  
  lims <- mh_data %>%
    group_by(CHR) %>%
    summarise(avg_coord = (min(x_coord) + max(x_coord)) / 2)
  
  mh_cols <- setNames(
    rep(x=c("#999999", "#555555"), length.out = 22),  # Gray/dark gray for alternating chromosomes
    levels(factor(lims$CHR))
  )
  
  mh_cols <- c(
    setNames(rep(c("#999999", "#555555"), length.out = 22), levels(factor(lims$CHR))),  # Gray/dark gray for chromosomes
    setNames(RColorBrewer::brewer.pal(3, "Set1"), unique(mh_data$y))   # Unique colors for `y`
  )
  
  gwis_mh_plt <- ggplot() +
    geom_point(data = mh_data,
               aes(x = x_coord, y = nlp, color = factor(color)), 
               size = 0.6, alpha = 1) +
    geom_hline(yintercept = -log10(gwis_bonferroni), linetype = "dashed", color = "black") + 
    scale_x_continuous(breaks = lims$avg_coord[c(1:16, 18, 20, 20, 22)], 
                       labels = c(1:16, 18, 20, 20, 22)) +
    scale_y_continuous(name = expression(-log[10](italic(p))*" GxBMI")) +
    scale_colour_manual(values = mh_cols, breaks = unique(mh_data$y),
                        name = "Biomarker") +
    guides(color = "none") +
    theme_minimal() +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_text(vjust = -1.5),
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_blank())
  
  gwis_mh_plt
}
```

## ALT

```{r alt-fig-full, fig.asp=1}
alt_base_pgs_plt <- generate_base_plot("alt_log", "log(ALT)")
alt_ppgs_plt <- generate_ppgs_plot("alt_log", "log(ALT)")
alt_gwis_mh_plt <- generate_gwis_plot("alt_log", "log(ALT)")

(
  alt_ppgs_plt /
    alt_gwis_mh_plt
) +
  plot_layout(heights = c(4, 3)) +
  plot_annotation(tag_levels = "a")
```

## AST

```{r ast-fig-full, fig.asp=1}
ast_base_pgs_plt <- generate_base_plot("ast_log", "log(AST)")
ast_ppgs_plt <- generate_ppgs_plot("ast_log", "log(AST)")
ast_gwis_mh_plt <- generate_gwis_plot("ast_log", "log(AST)")

(
  ast_ppgs_plt /
    ast_gwis_mh_plt
) +
  plot_layout(heights = c(4, 3)) +
  plot_annotation(tag_levels = "a")
```

## GGT

```{r ggt-fig-full, fig.asp=1}
ggt_base_pgs_plt <- generate_base_plot("ggt_log", "log(GGT)")
ggt_ppgs_plt <- generate_ppgs_plot("ggt_log", "log(GGT)")
ggt_gwis_mh_plt <- generate_gwis_plot("ggt_log", "log(GGT)")

(
  ggt_ppgs_plt /
    ggt_gwis_mh_plt
) +
  plot_layout(heights = c(4, 3)) +
  plot_annotation(tag_levels = "a")
```

<!--
# Archive

```{r pgs-testing-prep, eval=F}
gPCs <- paste0("gPC", 1:10)
primary_covars <- c("sex", "age", "age_squared", "ageBySex", gPCs)
liver_bms <- c("alt_log", "ast_log", "ggt_log")
```

```{r ppgs-results-old, fig.asp=1.2, eval=F}
read_pgs_prset <- function(tag, pg) {
ppgs_df <- read_delim(paste0("../data/processed/pgs/", tag, "_prset_", pg, ".all_score"),
delim = " ", show_col_types = FALSE) %>%
select(id = IID, everything(), -FID)
# rename_with(~ str_replace_all(., "_[\\d\\.]+$", ""), -id)
ppgs_df
}

test_single_ppgs_gxe <- function(e, y, regression_df, covars, standardize) {
  if (standardize) {
    regression_df <- mutate(regression_df,
                            across(all_of(c(e, y, "ppgs")), ~ as.vector(scale(.x))))
  }
  lm_form_str <- paste0(y, " ~ ppgs * ", e)
  if (!is.null(covars)) {
    lm_form_str <- paste0(lm_form_str, " + ", paste(covars, collapse = " + "))
  }
  lm_fit <- lm(as.formula(lm_form_str), data = regression_df)
  lm_fit %>%
    broom::tidy() %>%
    filter(term == paste0("ppgs:", e))
}

test_all_ppgs_gxe <- function(e, y, pg, threshold, set, covars, 
                          ancestry_test = "all", standardize = TRUE) {
  ppgs_df <- read_pgs_prset(y, pg) %>%
    select(id, contains(threshold))
  pathways <- gsub(paste0("_", threshold), "", setdiff(names(ppgs_df), "id"))
  regression_df <- inner_join(ukb_subsets[[set]], ppgs_df, by = "id")
  if (ancestry_test != "all") {
    regression_df <- inner_join(regression_df, ancestry_df, by = "id") %>%
      filter(ancestry == ancestry_test)
  }
  map(pathways[1:4], function(p) {
    regression_df$ppgs <- regression_df[[paste0(p, "_", threshold)]]
    test_single_ppgs_gxe(e, y, regression_df, covars, standardize)
  }) %>%
    setNames(pathways[1:4]) %>%
    bind_rows(.id = "pathway")
}
# test_all_ppgs_gxe("bmi", "alt_log", "0.001", "testing",
#                   c(primary_covars, paste0("bmiBy", gPCs)))

ppgs_res_df <- ppgs_config_df %>%
  select(e, y, pathway_group, threshold) %>%
  rowwise() %>%
  mutate(pathway_fits = list(
    test_all_ppgs_gxe(e, y, pathway_group, threshold, "testing",
                      c(primary_covars, paste0(e, "By", gPCs)))
  )) %>%
  unnest(pathway_fits)
```

```{r, eval=F}
locus_to_pathway_df %>%
  mutate(pathway = gsub("HALLMARK_|KEGG_|KEGG_MEDICUS_", "", pathway)) %>%
  left_join(
    filter(ppgs_res_df, y == "alt_log", threshold == 0.001) %>%
      select(pathway, pathway_p = p.value), 
    by = c("pathway")) %>% 
  arrange(CHR, POS, pathway_p) %>%
  filter(pathway_group == "kegg_legacy") %>%
  kable(caption = "ALT-specific SNP to pathway mapping") %>%
  kable_styling()
```

```{r snp-to-gene-mapping-old, eval=F}
manual_snp_to_gene_map_df <- tribble(
  ~rsid, ~gene, ~join_gene,
  "rs2642438", "MTARC1", "MTARC2",
  "rs17036160", "PPARG", "PPARG",
  "rs11735092", "HSD17B13", "HSD17B11",
  "rs2954021", "TRIB1", "TRIB1",
  "rs5787948", "GPAM", "GPAM",
  "rs200210321", "SUGP1/TM6SF2", "TM7SF2",
  "rs429358", "APOE", "APOE",
  "rs3747207", "PNPLA3", "PNPLA3"
)

locus_to_pathway_df <- pruned_gwis_ss$alt_log %>%
  select(index_var, CHR, POS, P_int) %>%
  left_join(manual_snp_to_gene_map_df, by = c("index_var" = "rsid")) %>%
  left_join(gene_pathway_annot_df, by = c("join_gene" = "gene"))

locus_to_pathway_df %>%
  write_csv("../output/alt_locus_to_pathway.csv")

locus_to_pathway_df %>%
  select(index_var, CHR, POS, gene, join_gene, pathway) %>%
  kable(caption = "GWIS locus to pathway mapping") %>%
  kable_styling(full_width = FALSE)
```
-->

# Export pathways of interest

```{r export-pathways}
# for (pwy_grp in pathway_groups) {
#   gene_pathway_annot_df %>%
#     filter(pathway_group == pwy_grp) %>%
#     distinct(pathway) %>%
#     pull(pathway) %>%
#     write(paste0("../data/processed/pgs/", pwy_grp, "_all_pwy_names.txt"))
# }

for (pwy_grp in pathway_groups) {
  for (outcome in outcomes) {
    sig_ppgs_res_df %>%
      filter(pathway_group == pwy_grp,
             y == outcome,
             threshold == 0.001,
             pathway != "Base") %>%
      distinct(pathway) %>%
      pull(pathway) %>%
      write(paste0("../data/processed/pgs/", outcome, "_", pwy_grp, "_sig_pwy_names.txt"))
  }
}
```
