---
output: html_document
title: "UKB results for ipgs proof-of-concept study"
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = F, message = F,warning = F, 
                      dpi = 150, fig.path = "../output/ukb_results/",
                      cache.path = "../cache/ukb_results/")
suppressMessages(silent <- lapply(
  c("knitr", "kableExtra", "tidyverse", "patchwork"), 
  library, character.only = TRUE))
theme_set(theme_bw())
```

```{r load-data}
ukb_subset_names <- c("training", "validation", "testing")
ukb_subset_names_clean <- c("Training", "Validation", "Testing")
ukb_subsets <- map(set_names(ukb_subset_names), function(set) {
  read_csv(paste0("../data/processed/ukb_", set, "_set.csv"), show_col_types = FALSE) %>%
    mutate(across(matches("^gPC"), ~ . * bmi, .names = "bmiBy{.col}"),
           across(matches("^gPC"), ~ . * whr, .names = "whrBy{.col}"))
})

ancestry_df <- read_csv("../data/processed/ukb_phenos_panUKBB.csv", 
                        col_select = c(id, ancestry), show_col_types = FALSE)
ancestries <- unique(ancestry_df$ancestry[!is.na(ancestry_df$ancestry)])
```

# Concept summary: pPGS for GxE

```{r workflow-diagram, eval=F}
knitr::include_graphics("../doc/workflow_diagram.pdf")
```

```{r configuration}
exposures <- c("bmi")
exposures_clean <- c("BMI")

# outcomes <- c("alb", "alt_log")
# outcomes_clean <- c("Albumin", "log(ALT)")
outcomes <- c("hscrp_log", "hba1c", "ldl_statinadj", "tg_log", "alt_log",
              "ggt_log", "hdl", "alb", "lipA_log", "vitD")
outcomes_clean <- c("log(hsCRP)", "HbA1c", "LDL-C", "log(TG)", "log(ALT)",
                    "log(GGT)", "HDL-C", "Albumin", "log(LipA)", "Vitamin D")

threshold_vec <- 5 * 10 ** seq(-8, -2)
thresholds <- c("5e-08", "0.005")
threshold_names <- paste0("Pt_", c("5e-08", "0.005"))
threshold_names_clean <- c("P<5e-8", "P<0.005")

pgs_config_df <- expand_grid(
  e = exposures,
  y = outcomes,
  set = names(ukb_subsets),
  threshold = c(thresholds, "all")
) %>%
  mutate(tag = paste0(y, "_prset"))
```

Overall dataset consists of unrelated, multi-ancestry individuals from the UKBB (based on Pan-UKBB; N ~ 320,000).

* 50% training set: GWAS
* 25% validation set: pPGS parameter optimization (pathway-level weights)
* 25% testing set: evaluation

Primary metric for evaluation (validation & testing): significance of regression PGSxE term

Variables being manipulated:

* Exposure: BMI
* Outcome: 20 cardiometabolic risk factors
* P&T threshold: 0.005, 5e-8

# Data distributions

Plots reflect the testing set.

```{r histograms, out.width="40%"}
plot_histogram <- function(f) {
  ukb_subsets$testing %>%
    select(x = {{ f }}) %>%
    filter(!is.na(x)) %>%
    ggplot(aes(x = x)) +
    geom_histogram(stat = "bin", bins = 30) +
    labs(x = f)
}

hscrp_hist <- plot_histogram("hscrp_log")
hba1c_hist <- plot_histogram("hba1c")

hscrp_hist + hba1c_hist

for (outcome in outcomes) {
  print(plot_histogram(outcome))
}
```

# PGS performance

```{r pgs-testing-prep}
# read_pgs <- function(tag) {
#   pgs_df <- read_csv(paste0("../data/processed/pgs/", tag, "_all_pgs.csv"), 
#                      show_col_types = FALSE)
#   pgs_df
# }
# 
# read_pgs_prsice <- function(tag) {
#   pgs_df <- read_delim(paste0("../data/processed/pgs/", tag, ".all_score"), 
#                        delim = " ", show_col_types = FALSE) %>%
#     select(-FID, id = IID, everything())
#   pgs_df
# }

# test_pgs_main <- function(y, tag, set, thresh_name, covars = NULL) {
#   pgs_df <- read_pgs_prsice(tag)
#   if (thresh_name == "stacked") {
#     stack_weights <- pull(filter(stack_weights_df, tag == .env$tag), stack_weights)[[1]]
#     pgs_df$stacked <- calc_stacked_pgs(pgs_df, stack_weights)
#   }
#   pgs_df <- select(pgs_df, id, pgs = {{thresh_name}})
#   regression_df <- inner_join(ukb_subsets[[set]], pgs_df, by = "id")
#   lm_form_str <- paste0(y, " ~ pgs")
#   if (!is.null(covars)) {
#     lm_form_str <- paste0(lm_form_str, " + ", paste(covars, collapse = " + "))
#   }
#   tryCatch({
#     lm_fit <- lm(as.formula(lm_form_str), data = regression_df)
#     lm_fit %>%
#       broom::tidy() %>%
#       filter(term == "pgs")
#   }, error = function(e) tibble())
# }

# test_pgs_gxe <- function(e, y, tag, set, thresh_name, covars = NULL) {
#   # pgs_df <- read_csv(paste0("../data/processed/pgs/", tag, "_all_pgs.csv"), 
#   #                    show_col_types = FALSE)
#   # if (thresh_name == "stacked") {
#   #   stack_weights <- pull(filter(stack_weights_df, tag == .env$tag), stack_weights)[[1]]
#   #   pgs_df$stacked <- calc_stacked_pgs(pgs_df, stack_weights)
#   # }
#   pgs_df <- tryCatch({
#     pgs_df <- read_pgs_prsice(tag)
#     pgs_df <- select(pgs_df, id, pgs = {{thresh_name}})
#     regression_df <- inner_join(ukb_subsets[[set]], pgs_df, by = "id")
#   }, error = function(e) tibble())
#   lm_form_str <- paste0(y, " ~ pgs * ", e)
#   if (!is.null(covars)) {
#     lm_form_str <- paste0(lm_form_str, " + ", paste(covars, collapse = " + "))
#   }
#   tryCatch({
#     lm_fit <- lm(as.formula(lm_form_str), data = regression_df)
#     lm_fit %>%
#       broom::tidy() %>%
#       filter(term == paste0("pgs:", e))
#   }, error = function(e) tibble())
# }

gPCs <- paste0("gPC", 1:10)
primary_covars <- c("sex", "age", "age_squared", "ageBySex", gPCs)
```

```{r mpgs-results, eval=F}
mpgs_test_res_df <- pgs_config_df %>%
  filter(pgs_type == "mpgs") %>%
  rowwise() %>%
  mutate(lm_fit = list(test_pgs_gxe(e, y, tag, set, thresh_name, 
                                    c(primary_covars, paste0(e, "By", gPCs))))) %>%
  ungroup() %>%
  unnest(lm_fit)

optimal_mpgs_df <- mpgs_test_res_df %>%
  filter(set == "validation") %>%
  group_by(e, y) %>%
  filter(statistic == max(statistic)) %>%
  ungroup() %>%
  mutate(optimal = TRUE) %>%
  select(e, y, pgs_type, thresh_name, optimal)

mpgs_test_res_df %>%
  filter(set == "validation") %>%
  left_join(optimal_mpgs_df, by = c("e", "y", "pgs_type", "thresh_name")) %>%
  mutate(y = factor(y, levels = outcomes, labels = outcomes_clean),
         thresh_name = factor(thresh_name, 
                              levels = threshold_names,
                              labels = threshold_names_clean)) %>%
  ggplot(aes(x = thresh_name, y = statistic, color = optimal)) +
  geom_bar(stat = "identity") +
  scale_color_discrete(name = "Optimal", na.translate = FALSE) +
  facet_wrap(vars(y, e), nrow = 2, scale = "free_y") +
  labs(x = "", y = "mPGS GxE z-statistic",
       title = "Optimization of mPGS for PGSxE") +
  theme(axis.text.x = element_text(angle = 25, hjust = 0.9))
```

```{r ppgs-results, fig.asp=1.2}
# read_pgs_prset <- function(tag, thresh) {
#   ppgs_df <- read_delim(paste0("../data/processed/pgs/", tag, ".all_score"), 
#                        delim = " ", show_col_types = FALSE) %>%
#     select(id = IID, everything(), -FID, -contains("Base"))
#     # rename_with(~ str_replace_all(., "_[\\d\\.]+$", ""), -id)
#   ppgs_df
# }

# get_ppgs_weights <- function(e, y, tag, thresh, set, covars) {
#   pgs_df <- read_pgs_prset(tag, thresh)
#   if (!(thresh == "all")) {
#     pgs_df <- select(pgs_df, id, contains(thresh))
#   }
#   pathways <- setdiff(names(pgs_df), "id")
#   regression_df <- inner_join(ukb_subsets[[set]], pgs_df, by = "id")
#   lm_form_str <- paste0(y, " ~ pgs * ", e)
#   if (!is.null(covars)) {
#     lm_form_str <- paste0(lm_form_str, " + ", paste(covars, collapse = " + "))
#   }
#   univariate_ppgs_gxe_df <- parallel::mclapply(
#     setNames(pathways, pathways), 
#     function(p) {
#       regression_df$pgs <- regression_df[[p]]
#       lm_fit <- lm(as.formula(lm_form_str), data = regression_df)
#       lm_fit %>%
#         broom::tidy() %>%
#         filter(term == paste0("pgs:", e))
#     }, 
#     mc.cores = 16) %>%
#     bind_rows(.id = "pathway")
#   univariate_ppgs_gxe_df
# }

ppgs_calc_df <- readRDS("../data/processed/pgs/ppgs_vec_df.rds")

test_ppgs_gxe <- function(e, y, tag, set, covars, ppgs_df, 
                          ancestry_test = "all", standardize = TRUE) {
  # pgs_df <- read_pgs_prset(tag)
  # # stopifnot(identical(names(pgs_df)[-1], weights_df$pathway))
  # pgs_df$ppgs <- as.matrix(pgs_df[weights_df$pathway]) %*% weights_df$estimate
  regression_df <- inner_join(ukb_subsets[[set]], ppgs_df, by = "id")
  if (ancestry_test != "all") {
    regression_df <- inner_join(regression_df, ancestry_df, by = "id") %>%
      filter(ancestry == ancestry_test)
  }
  if (standardize) {
    regression_df <- mutate(regression_df, 
                            across(all_of(c(e, y, "ppgs")), ~ as.vector(scale(.x))))
  }
  names(regression_df)[length(regression_df)] <- "ppgs"  ### TEMPORARY
  lm_form_str <- paste0(y, " ~ ppgs * ", e)
  if (!is.null(covars)) {
    lm_form_str <- paste0(lm_form_str, " + ", paste(covars, collapse = " + "))
  }
  lm_fit <- lm(as.formula(lm_form_str), data = regression_df)
  lm_fit %>%
    broom::tidy() %>%
    filter(term == paste0("ppgs:", e))
}

# ppgs_weights_df <- pgs_config_df %>%
#   select(e, y, tag, threshold) %>%
#   distinct() %>% 
#   rowwise() %>%
#   mutate(pathway_fits = list(
#     get_ppgs_weights(e, y, paste0(y, "_prset"), threshold, "validation", 
#                      c(primary_covars, paste0(e, "By", gPCs)))
#   ))

ppgs_test_res_df <- ppgs_calc_df %>%
  expand_grid(set = names(ukb_subsets),
              ancestry = c("all", "EUR")) %>%
  # inner_join(ppgs_weights_df, by = c("e", "y", "tag", "threshold")) %>%
  rowwise() %>%
  mutate(lm_fit = list(test_ppgs_gxe(e, y, tag, set, 
                                     c(primary_covars, paste0(e, "By", gPCs)),
                                     ppgs_df, ancestry_test = ancestry))) %>%
  ungroup() %>%
  unnest(lm_fit)
```

```{r, fig.asp=1.2}
ppgs_calc_df %>%
  unnest(pathway_fits) %>%
  arrange(estimate) %>%
  mutate(pathway = factor(pathway, levels = pathway, 
                          labels = str_replace(pathway, "HALLMARK_", ""))) %>%
  ggplot(aes(x = pathway, y = estimate)) +
  geom_bar(stat = "identity") +
  facet_wrap(vars(e, y), ncol = 2, scale = "free_y") +
  labs(x = "", y = "pPGSxE effect size (final pPGS weight)",
       title = "Optimization of pPGS for PGSxE (validation set)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9, size = 4))
```

```{r plot-pgs-results, fig.asp=1.2}
n_biomarkers <- 20
bonferroni <- 0.05 / n_biomarkers

ppgs_test_res_df %>%
  filter(set == "testing") %>%
  mutate(e = factor(e, levels = exposures, labels = exposures_clean),
         y = factor(y, levels = outcomes, labels = outcomes_clean),
         # pgs_type = factor(pgs_type, levels = pgs_types, labels = pgs_types_clean),
         set = factor(set, levels = names(ukb_subsets), 
                      labels = paste0(str_to_title(names(ukb_subsets)), " set"))) %>%
         # thresh_name = factor(thresh_name, 
         #                      levels = threshold_names,
         #                      labels = threshold_names_clean)) %>%
  ggplot(aes(x = threshold, y = statistic)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 0, color = "gray") +
  geom_hline(yintercept = qnorm(0.05 / 2, lower.tail = FALSE), linetype = "dashed") +
  geom_hline(yintercept = qnorm(bonferroni / 2, lower.tail = FALSE), linetype = "dashed") +
  scale_fill_discrete(name = "PGS type") +
  facet_wrap(vars(y), ncol = 2, scale = "fixed") +
  labs(x = "", y = "Interaction test z-statistic",
       title = "PGSxE testing (testing set)") +
  theme(axis.text.x = element_text(angle = 20, hjust = 0.9))

ppgs_test_res_df %>%
  filter(set == "testing") %>%
  group_by(y) %>%
  summarise(any_nom = any(p.value < 0.05),
            any_bonf = any(p.value < bonferroni)) %>%
  count(any_nom, any_bonf) %>%
  arrange(desc(n))

# all_pgs_test_res_df %>%
#   filter(set == "testing") %>%
#   group_by(y) %>%
#   summarise(top_pgs_type = pgs_type[which.max(statistic)]) %>%
#   count(top_pgs_type) %>%
#   arrange(desc(n))

ppgs_test_res_df %>%
  filter(set == "testing") %>%
  arrange(p.value) %>%
  head(5)
```

```{r, fig.asp=1.2}
pgs_types <- c("mpgs", "ppgs", "ipgs", "vpgs")
pgs_types_clean <- c("mPGS", "pPGS", "iPGS", "vPGS")

multi_pgs_test_res_df <- readRDS("../../ipgs/data/processed/manuscript/all_ukb_pgs_test_res_df.rds") %>%
  filter(set == "testing") %>%
  select(e, y, pgs_type, estimate, std.error, statistic, p.value)

all_pgs_test_res_df <- bind_rows(list(
  ppgs_test_res_df %>% 
    mutate(pgs_type = "ppgs") %>% 
    filter(threshold == "all", ancestry == "EUR", set == "testing"),
  multi_pgs_test_res_df
))

all_pgs_test_res_df %>%
  mutate(e = factor(e, levels = exposures, labels = exposures_clean),
         y = factor(y, levels = outcomes, labels = outcomes_clean),
         pgs_type = factor(pgs_type, levels = pgs_types, labels = pgs_types_clean),
         set = factor(set, levels = names(ukb_subsets), 
                      labels = paste0(str_to_title(names(ukb_subsets)), " set")),
         l95 = estimate - 1.96 * std.error,
         u95 = estimate + 1.96 * std.error) %>%
         # thresh_name = factor(thresh_name, 
         #                      levels = threshold_names,
         #                      labels = threshold_names_clean)) %>%
  filter(!is.na(y)) %>%
  ggplot(aes(x = pgs_type, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = l95, ymax = u95), width = 0) +
  geom_hline(yintercept = 0, color = "gray") +
  # geom_hline(yintercept = qnorm(0.05 / 2, lower.tail = FALSE), linetype = "dashed") +
  # geom_hline(yintercept = qnorm(bonferroni / 2, lower.tail = FALSE), linetype = "dashed") +
  scale_fill_discrete(name = "PGS type") +
  facet_wrap(vars(y), ncol = 2, scale = "fixed") +
  labs(x = "", y = "Interaction test z-statistic",
       title = "PGSxE testing (testing set)") +
  theme(axis.text.x = element_text(angle = 20, hjust = 0.9))
```

```{r}
ppgs_test_res_df %>%
  filter(set == "testing") %>%
  select(tag, threshold, statistic, ancestry) %>%
  pivot_wider(names_from = "ancestry", values_from = "statistic") %>%
  ggplot(aes(x = all, y = EUR)) +
  geom_point() +
  geom_abline()
```

# Specific visualizations

```{r example-viz, eval=F}
e <- "bmi"
y <- "hscrp_log"
tag <- "bmi_hscrp_log"
thresh_name <- "Pt_5e-08"
e_clean <- "BMI"
y_clean <- "log(hsCRP)"
# ppgs_df <- filter(ppgs_df, e == e)
regression_df <- inner_join(ukb_subsets$testing, ppgs_df, by = "id") %>%
  filter(!is.na(.data[[e]]), !is.na(.data[[y]])) %>%
  mutate(pgs_bin = cut(pgs, quantile(pgs, seq(0, 1, length.out = 11)), 
                       include.lowest = TRUE, labels = paste0("Q", 1:10)))
```

```{r, eval = FALSE}
regression_df %>%
  ggplot(aes(x = bmi, y = hscrp_log)) +
  geom_smooth(method = "gam", formula = y ~ rms::rcs(x, 5)) +
  labs(x = e_clean,
       y = y_clean)
```

```{r, eval = FALSE}
for (tag in c("hscrp_log", "bmi_hscrp_log", "hscrp_log_vQTL", "hscrp_log_prset")) {
  df <- read_table(paste0("../data/processed/pgs/", tag, ".all_score"), 
                 show_col_types = FALSE)
  print(head(df))
}
```

```{r, eval = FALSE}
pgs_test_main_res_df %>%
  left_join(optimal_pgs_main_df, by = c("y", "pgs_type", "thresh_name")) %>%
  filter(y == "hscrp_log") %>%
  mutate(color = if_else(set == "validation", optimal, NA),
         statistic = if_else(set == "testing" & optimal == FALSE, NA, statistic),
         y = factor(y, levels = outcomes, labels = paste0(outcomes_clean, " set")),
         set = factor(set, levels = ukb_subset_names, 
                      labels = ukb_subset_names_clean),
         thresh_name = factor(thresh_name, 
                              levels = threshold_names,
                              labels = threshold_names_clean)) %>%
  ggplot(aes(x = thresh_name, y = statistic, color = color)) +
  geom_bar(stat = "identity") +
  scale_color_discrete(name = "Optimal", na.translate = FALSE) +
  facet_wrap(vars(set), nrow = 1, scale = "free_y") +
  labs(x = "", y = "PGS main effect z-statistic",
       title = "Positive control: standard GWAS/main effects pipeline") +
  theme(axis.text.x = element_text(angle = 25, hjust = 0.9))
```

```{r, fig.width=5, eval = FALSE}
all_pgs_test_res_df %>%
  filter(set == "testing",
         e == "bmi",
         y == "hscrp_log") %>%
  mutate(e = factor(e, levels = exposures, labels = exposures_clean),
         y = factor(y, levels = outcomes, labels = outcomes_clean),
         pgs_type = factor(pgs_type, levels = pgs_types, labels = pgs_types_clean),
         set = factor(set, levels = names(ukb_subsets), 
                      labels = paste0(str_to_title(names(ukb_subsets)), " set")),
         thresh_name = factor(thresh_name, 
                              levels = threshold_names,
                              labels = threshold_names_clean)) %>%
  ggplot(aes(x = pgs_type, y = statistic, fill = pgs_type)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 0, color = "gray") +
  scale_fill_discrete(name = "PGS type") +
  labs(x = "", y = "Interaction test z-statistic",
       title = "PGSxE testing (testing set)") +
  theme(axis.text.x = element_text(angle = 20, hjust = 0.9))
```

```{r, eval = FALSE}
quantile_lines_plt <- regression_df %>%
  ggplot(aes(x = bmi, y = hscrp_log, group = pgs_bin, color = pgs_bin)) +
  # ggplot(aes(x = {{e}}, y = {{y}}, group = pgs_bin, color = pgs_bin)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method = "lm", formula = "y ~ x") +
  scale_color_discrete(name = "PGS bin") +
  coord_cartesian(xlim = quantile(regression_df[[e]], c(0.001, 0.999))) +
  labs(x = e_clean,
       y = y_clean)

test_ey <- function(df, e, y, covars) {
  lm_form_str <- paste0(y, " ~ ", e, " + ", paste(covars, collapse = " + "))
  lm_fit <- lm(as.formula(lm_form_str), data = df)
  lm_fit %>%
    broom::tidy() %>%
    filter(term == e)
}

quantile_effects_df <- regression_df %>%
  nest(data = -pgs_bin) %>%
  rowwise() %>%
  mutate(ey_lm = test_ey(data, e, y, primary_covars)) %>%
  select(-data) %>%
  unnest(ey_lm)
quantile_effects_plt <- quantile_effects_df %>%
  select(pgs_bin, estimate) %>%
  ggplot(aes(x = pgs_bin, y = estimate)) +
  geom_point() +
  labs(x = "PGS bin",
       y = paste0(e_clean, " - ", y_clean, " effect size"))

quantile_lines_plt + quantile_effects_plt

quantile_effects_df %>%
  arrange(pgs_bin) %>%
  select(pgs_bin, estimate) %>%
  mutate(factor_change = estimate / estimate[1])

# q1_corr <- cor.test(~ hscrp_log + whr, data = filter(regression_df, pgs_bin == "Q1"))$estimate
# q10_corr <- cor.test(~ hscrp_log + whr, data = filter(regression_df, pgs_bin == "Q10"))$estimate
# 
# print(paste0("WHR-log(hsCRP) correlation in lowest vs. highest decile: ",
#              round(q1_corr, 2), " versus ", round(q10_corr, 2)))
```

```{r, fig.asp=1.4, eval = FALSE}
all_pgs_test_res_df %>%
  filter(set == "testing",
         e == "bmi") %>%
  mutate(e = factor(e, levels = exposures, labels = exposures_clean),
         y = factor(y, levels = outcomes, labels = outcomes_clean),
         pgs_type = factor(pgs_type, levels = pgs_types, labels = pgs_types_clean),
         set = factor(set, levels = names(ukb_subsets), 
                      labels = paste0(str_to_title(names(ukb_subsets)), " set")),
         thresh_name = factor(thresh_name, 
                              levels = threshold_names,
                              labels = threshold_names_clean)) %>%
  ggplot(aes(x = pgs_type, y = statistic, fill = pgs_type)) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 0, color = "gray") +
  scale_fill_discrete(name = "PGS type") +
  facet_wrap(vars(y), ncol = 1, scale = "free_y") +
  labs(x = "", y = "Interaction test z-statistic",
       title = "iPGSxBMI results (testing set)") +
  theme(axis.text.x = element_text(angle = 20, hjust = 0.9))
```

```{r, eval = FALSE}
ppgs_weights_df %>%
  filter(e == "bmi",
         y == "hscrp_log") %>%
  unnest(pathway_fits) %>%
  arrange(estimate) %>%
  mutate(pathway = factor(pathway, levels = pathway, 
                          labels = str_replace(pathway, "HALLMARK_", ""))) %>%
  ggplot(aes(x = pathway, y = estimate)) +
  geom_bar(stat = "identity") +
  labs(x = "", y = "pPGSxBMI univariate effect size",
       title = "Example pPGS optimization (validation set)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.9))
```

# Export PGS definitions

```{r export-optimized-params, eval = FALSE}
all_pgs_test_res_df %>%
  mutate(threshold = threshold_vec[match(thresh_name, threshold_names)]) %>%
  filter(pgs_type == "mpgs",
         e == "bmi") %>%
  distinct(tag, threshold) %>%
  mutate(threshold = ifelse(threshold == 0.05, 0.005, threshold)) %>%  # REVISIT THIS *******
  write_csv("../data/processed/pgs/optimized_mpgs_params.csv")

all_pgs_test_res_df %>%
  mutate(threshold = threshold_vec[match(thresh_name, threshold_names)]) %>%
  filter(pgs_type == "ipgs",
         e == "bmi") %>%
  distinct(tag, threshold) %>%
  write_csv("../data/processed/pgs/optimized_ipgs_params.csv")

all_pgs_test_res_df %>%
  mutate(threshold = threshold_vec[match(thresh_name, threshold_names)]) %>%
  filter(pgs_type == "vpgs",
         e == "bmi") %>%
  distinct(tag, threshold) %>%
  write_csv("../data/processed/pgs/optimized_vpgs_params.csv")

all_pgs_test_res_df %>%
  filter(pgs_type == "ppgs",
         e == "bmi") %>%
  distinct(tag, pathway_fits) %>%
  unnest(pathway_fits) %>%
  select(tag, pathway, estimate) %>%
  pivot_wider(names_from = "pathway", values_from = "estimate") %>%
  write_csv("../data/processed/pgs/optimized_ppgs_params.csv")
```

```{r exploration, eval = FALSE}
e <- "whr"
y <- "hscrp_log"
tag <- "whr_hscrp_log"
thresh_name <- "thresh0.01"
covars <- NULL
covars <- primary_covars
covars <- c(primary_covars, paste0("bmi", "By", gPCs))

pgs_df <- read_csv(paste0("../data/processed/pgs/", tag, "_all_pgs.csv"), 
                   show_col_types = FALSE)
pgs_df <- select(pgs_df, id, pgs = {{thresh_name}})

set <- "training"
regression_df <- inner_join(ukb_subsets[[set]], pgs_df, by = "id")
lm_form_str <- paste0(y, " ~ pgs * ", e)
if (!is.null(covars)) {
  lm_form_str <- paste0(lm_form_str, " + ", paste(covars, collapse = " + "))
}
lm_fit <- lm(as.formula(lm_form_str), data = regression_df %>% sample_n(70000))
lm_fit %>%
  broom::tidy() %>%
  filter(term == paste0("pgs:", e))
```
